Sequence Divergence as the main modulator of structure-sequence correlations

### 2014-10-22 manuscript/cordiv.tex
MN revisions

        For the measure of local packing density in proteins (the Weighted Contact Number) we find that the average of coordinates of all heavy atoms for each residue results in WCN values that show the best correlation with other structural properties.

>>>>

        For the measure of local packing density in proteins (the Weighted Contact Number) we find that among all possible set of coordinates, the average over coordinates of all heavy atoms for each each individual side chain results in WCN values that show the best correlation with other structural and sequence properties, such as RSA, Voronoi cell properties, sequence entropy, and evolutionary rates. Specifically, WCN from average side chain coordinates (wcnSC) outperforms WCN based on $C_\alpha$ coordinates (wcnCA) in predicting RSA, Voronoi cell volumes, sequence entropy and evolutionary rates by a median Spearman correlation difference of $00$, $00$, $00$ \& $00$, respectively (Figure \ref{fig:best_wcn}

        For the measure of local flexibility in proteins (Bfactor) we similarly find that among all $7$ representative measures of site Bfactors, the average of Bfactor values over all heavy atoms of each individual side chain (bfSC) results in the best correlations with other structural and sequence properties. Specifically, bfSC outperforms the commonly used $C_\alpha$ Bfactor (bfCA) in predicting RSA, Voronoi cell volumes, sequence entropy and evolutionary rates by a median Spearman correlation difference of $00$, $00$, $00$ \& $00$, respectively.

        Similar to WCN and Bfactor, the Voronoi cell properties, most importantly the cell surface area, volume and the cell compactness also correlate best with other structure and sequence properties, only if the average side chain coordinates are used as the seeds of Voronoi cells.

    \begin{figure}[tbh]
        \begin{center}
        \includegraphics[width=6.9in]{/best_wcn/selected_variables/boxplot_wcn_all_in_one.pdf} % Fig. 1
        \end{center}
        \caption{A comparison of the strength of Spearman correlation of sequence evolutionary rates (r4sJC) with {\it side chain} Weighted Contact Number vs. correlations of other structural properties with evolutionary rates. Detailed description of the structural properties is given in Section \label{sec:mam}. The red lines in each plot represent equality line. It is evident from all plots that for any given protein in dataset, the correlation strength of one structural property is a good proxy measure of the correlation strength of any other structural property with sequence variability measures. For brevity, correlations of structure-rate4site are not shown here but are available online, also in supplementary material.}
        \label{fig:best_wcn}
    \end{figure}

    \begin{figure}[tbh]
        \begin{center}
        \includegraphics[width=6.9in]{/best_bf/selected_variables/boxplot_bf_all_in_one.pdf} % Fig. 1
        \end{center}
        \caption{A comparison of the strength of Spearman correlation of sequence evolutionary rates (r4sJC) with {\it side chain} Weighted Contact Number vs. correlations of other structural properties with evolutionary rates. Detailed description of the structural properties is given in Section \label{sec:mam}. The red lines in each plot represent equality line. It is evident from all plots that for any given protein in dataset, the correlation strength of one structural property is a good proxy measure of the correlation strength of any other structural property with sequence variability measures. For brevity, correlations of structure-rate4site are not shown here but are available online, also in supplementary material.}
        \label{fig:best_wcn}
    \end{figure}

    \begin{figure}[tbh]
        \begin{center}
        \includegraphics[width=6.9in]{/best_vorvol/selected_variables/boxplot_vorvol_all_in_one.pdf} % Fig. 1
        \end{center}
        \caption{A comparison of the strength of Spearman correlation of sequence evolutionary rates (r4sJC) with {\it side chain} Weighted Contact Number vs. correlations of other structural properties with evolutionary rates. Detailed description of the structural properties is given in Section \label{sec:mam}. The red lines in each plot represent equality line. It is evident from all plots that for any given protein in dataset, the correlation strength of one structural property is a good proxy measure of the correlation strength of any other structural property with sequence variability measures. For brevity, correlations of structure-rate4site are not shown here but are available online, also in supplementary material.}
        \label{fig:best_wcn}
    \end{figure}
